Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-24 numerical-gravity.tex
Section 5 edits (small)

First, the massive (dark matter and star) particles are distributed onto the grids using the second-order cloud-in-cell (CIC) interpolation technique \citep{Hockney88} to form a spatially-discretized density field $\rho_{\rm DM}$.  During the CIC interpolation, particle positions are (temporarily) advanced by $0.5 \Delta t v^n$ so that we generate an estimate of the time-centered density field.  Particles on subgrids within the grid's volume are also added to its gravitating field using the same method. In addition, since the gravitating field for a grid is defined beyond the grid edges (see below), massive particles from sibling grids and sub grids that lie within the entire gravitating field are used.  This step can involve communication.

Next, we add the baryonic grid densities in a similar fashion.  In particular, we treat baryonic cells as virtual CIC particles that are are placed at the grid center but are advanced by $0.5 \Delta t v^n$ in order to approximately time-center the gravitating mass field.  Cells that are covered by further-refined grids are treated in a similar way (i.e., we also use the subgrid cells as virtual CIC particles).  This procedure results in a total gravitating mass field $\rho_{\rm total}^{n+1/2}$.

>>>>

First, the massive (dark matter and star) particles are distributed onto the grids using the second-order cloud-in-cell (CIC) interpolation technique \citep{Hockney88} to form a spatially-discretized density field $\rho_{\rm DM}$.  During the CIC interpolation, particle positions are (temporarily) advanced by $0.5 v^n \Delta t$ so that we generate an estimate of the time-centered density field.  Particles on subgrids within the grid's volume are also added to its gravitating field using the same method. In addition, since the gravitating field for a grid is defined beyond the grid edges (see below), massive particles from sibling grids and sub grids that lie within the entire gravitating field are used.  This step can involve communication.

Next, we add the baryonic grid densities in a similar fashion.  In particular, we treat baryonic cells as virtual CIC particles that are are placed at the grid center but are advanced by $0.5 v^n \Delta t$ in order to approximately time-center the gravitating mass field.  Cells that are covered by further-refined grids are treated in a similar way (i.e., we also use the subgrid cells as virtual CIC particles).  This procedure results in a total gravitating mass field $\rho_{\rm total}^{n+1/2}$.

### 2013-06-24 numerical-particle.tex
Section 5 edits (small)

\enzo\ uses a particle-mesh N-body method to calculate the dynamics of
collisionless systems \citep{Hockney88}.  This method follows
trajectories of a representative sample of individual particles that
sample the phase space of the dark matter distribution, and is much
more efficient than a direct solution of the Boltzmann equation in
essentially all astrophysical situations for the levels of accuracy
that are required for simulations of structure formation.  As
described earlier, the gravitational potential is computed by solving
the elliptic Poisson's equation (Eq.~\ref{eq:potential}), and
differencing the potential to find accelerations, which are then
interpolated back to particles.  This acceleration is time-centered
(because the underlying gravitating mass field is approximately
time-centered), and so we have accelerations $\myvec{g}^{n+1/2}$ for
each particle.  These are used to update the particle positions and
velocities starting from $\myvec{x}^n$ and $\myvec{v}^n$ using a
standard drift-kick-drift technique \citep{Hockney88}:

@@

To avoid unphysical point-mass effects, \enzo\ provides a parameter that
governs the maximum level at which particles will be regarded as point masses.
At higher levels, contributions from particles to the gravitating mass field
will be smoothed over a spherical region centered at each particle's center.

>>>>

\enzo\ uses a particle-mesh N-body method to calculate the dynamics of
collisionless systems \citep{Hockney88}.  This method follows
trajectories of a representative sample of individual particles that
sample the phase space of the dark matter distribution, and is much
more efficient than a direct solution of the Boltzmann equation in
essentially all astrophysical situations for the levels of accuracy
that are required for simulations of structure formation.  As
described earlier, the gravitational potential is computed by solving
the elliptic Poisson's equation (Eq.~\ref{eq:potential}) and
differencing the potential to find accelerations, which are then
interpolated back to particles.  This acceleration is time-centered
(because the underlying gravitating mass field is approximately
time-centered), and so we have accelerations $\myvec{g}^{n+1/2}$ for
each particle.  These are used to update the particle positions and
velocities starting from $\myvec{x}^n$ and $\myvec{v}^n$ using a
standard drift-kick-drift technique \citep{Hockney88}:

@@

To avoid unphysical point-mass effects, \enzo\ provides a parameter that
governs the maximum level at which particles will be regarded as point masses.
At higher levels, contributions from particles to the gravitating mass field
will be smoothed over a spherical region centered at each particle's position.
